---
title: "wk7"
output: html_document
date: "2023-12-07"
---

## Topic: Spatial Autocorrelation
It is measure of similarity between nearby data.
Now we will load the packages.

```{r}
library(here)
library(janitor)
library(tidyverse)
library(sf)
```
load the data 
#Londonwards
```{r}
londonwards <- st_read(here::here("wk5data","statistical-gis-boundaries-london","statistical-gis-boundaries-london","ESRI","London_Ward.shp"))
```
read the londonward merged data
```{r}
londonWardsMerged <- st_read(here::here("wk5data","statistical-gis-boundaries-london","statistical-gis-boundaries-london","ESRI","London_Ward_CityMerged.shp"))%>%
  st_transform(.,27700)
```
read the csv file

```{r}
WardData <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv",
                    locale = locale(encoding = "latin1"),
                     na = c("NA", "n/a")) %>% 
  clean_names()
```
Now we want to join the shape file with the csv file.
we will do left join whch will not alter the left columns of the sahpe file.
```{r}
londonWardsMerged <- londonWardsMerged %>%
  left_join(WardData,
            by = c("GSS_CODE" ="new_code"))%>%
   dplyr::distinct(GSS_CODE, .keep_all = T)%>%
  dplyr::select(GSS_CODE, ward_name, average_gcse_capped_point_scores_2014)
```

Check whether the projection is right
```{r}
st_crs(londonWardsMerged)
```
Plot the map to check whether all the points are inside and the map not outside the region.

```{r}
library(tmap)

BluePlaques <- st_read(here::here("wk6data",
                                  "open-plaques-london-2018-04-08.geojson")) %>%
    st_transform(.,27700)
```
```{r}
tmap_mode("plot")
tm_shape(londonWardsMerged) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(BluePlaques) +
  tm_dots(col = "blue")
```
After plotting we find some blue plaques fall outside London so we need to remove it before doing any analysis.
##Data Cleaning
```{r}
summary(BluePlaques)
```
make a spatial subset which will remove the plaques outside london.
```{r}
BluePlaquesSub <- BluePlaques[londonWardsMerged,]
```

```{r}
tm_shape(londonWardsMerged) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(BluePlaquesSub) +
  tm_dots(col = "blue")
```
##Data manipulation

```{r}
example<-st_intersects(londonWardsMerged, BluePlaquesSub)

example
```
```{r}
check_example <- londonWardsMerged%>%
  st_join(BluePlaquesSub)%>%
  filter(ward_name=="Kingston upon Thames - Coombe Hill")
```

```{r}
library(sf)
points_sf_joined <- londonWardsMerged%>%
  mutate(n = lengths(st_intersects(., BluePlaquesSub)))%>%
  janitor::clean_names()%>%
  #calculate area
  mutate(area=st_area(.))%>%
  #then density of the points per ward
  mutate(density=n/area)%>%
  #select density and some other variables 
  dplyr::select(density, ward_name, gss_code, n, average_gcse_capped_point_scores_2014)
```

Make a Map
```{r}
points_sf_joined<- points_sf_joined %>%                    
  group_by(gss_code) %>%         
  summarise(density = first(density),
          wardname= first(ward_name),
          plaquecount= first(n))

tm_shape(points_sf_joined) +
    tm_polygons("density",
        style="jenks",
        palette="PuOr",
        midpoint=NA,
        popup.vars=c("wardname", "density"),
        title="Blue Plaque Density")
```
So, from the map, it looks as though we might have some clustering of blue plaques in the centre of London so let’s check this with Moran’s I and some other statistics.

#Weight Matrix
```{r}
library(spdep)
```
```{r}
#First calculate the centroids of all Wards in London

coordsW <- points_sf_joined%>%
  st_centroid()%>%
  st_geometry()
  
plot(coordsW,axes=TRUE)
```
```{r}
#create a neighbours list
LWard_nb <- points_sf_joined %>%
  poly2nb(., queen=T)
```

```{r}
summary(LWard_nb)
```
```{r}
#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
#add a map underneath
plot(points_sf_joined$geometry, add=T)
```
#Matrix Style

```{r}
#create a spatial weights matrix from these weights
Lward.lw <- LWard_nb %>%
  nb2mat(., style="B")

sum(Lward.lw)
```
```{r}
sum(Lward.lw[1,])
```
Standardisation permits comparable spatial parameters. But you also need to consider geographical features — should a neighbour be a neighbour if there is a river between them?

## Autocorrelation
Now we have defined our Wij matrix, we can calculate the Moran’s I and other associated statistics. However, Moran’s I requires a spatial weight list type object as opposed to matrix, this is simply…
```{r}
Lward.lw <- LWard_nb %>%
  nb2listw(., style="C")
```
# Moran's I
Moran’s I test tells us whether we have clustered values (close to 1) or dispersed values (close to -1), we will calculate for the densities rather than raw values (think about why this is?).

```{r}
I_LWard_Global_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  moran.test(., Lward.lw)

I_LWard_Global_Density
```
#Geary's C
 This tells us whether similar values or dissimilar values are clustering
```{r}
C_LWard_Global_Density <- 
  points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  geary.test(., Lward.lw)

C_LWard_Global_Density
```
#Getis Ord
This tells us whether high or low values are clustering. If G > Expected = High values clustering; if G < expected = low values clustering

```{r}
G_LWard_Global_Density <- 
  points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  globalG.test(., Lward.lw)

G_LWard_Global_Density
```
So the global statistics are indicating that we have spatial autocorrelation of Blue Plaques in London:  it suggests that the distribution of Blue Plaques is not random across the geography of London, but rather, there are spatial patterns or clusters.

##SUMMARY:

The Moran’s I statistic = 0.67 (remember 1 = clustered, 0 = no pattern, -1 = dispersed) which shows that we have some distinctive clustering

The Geary’s C statistic = 0.41 (remember Geary’s C falls between 0 and 2; 1 means no spatial autocorrelation, <1 - positive spatial autocorrelation or similar values clustering, >1 - negative spatial autocorreation or dissimilar values clustering) which shows that similar values are clustering

The General G statistic = G > expected, so high values are tending to cluster.

#Local Moran's I

```{r}
#use the localmoran function to generate I for each ward in the city

I_LWard_Local_count <- points_sf_joined %>%
  pull(plaquecount) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()

I_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()

#what does the output (the localMoran object) look like?
slice_head(I_LWard_Local_Density, n=5)
```
```{r}
points_sf_joined <- points_sf_joined %>%
  mutate(plaque_count_I = as.numeric(I_LWard_Local_count$Ii))%>%
  mutate(plaque_count_Iz =as.numeric(I_LWard_Local_count$Z.Ii))%>%
  mutate(density_I =as.numeric(I_LWard_Local_Density$Ii))%>%
  mutate(density_Iz =as.numeric(I_LWard_Local_Density$Z.Ii))
```

 


